Experiments and Modeling of the Autoignition of Methyl Pentanoate at Low to Intermediate Temperatures and Elevated Pressures in a Rapid Compression Machine

### 2017-09-25 revision-1/response/response.tex

\begin{response}
    %
    This statement has been removed from the manuscript.
    %
\end{response}

\begin{reviewer}
    %
    On discussion related to RMG model from Figure 5, the authors state " model over-predicts the
    low-temperature ignition delays and does not predict.." define low temperature ignition delays."
    The ignition delays from RMG model and experiments cross over around 720 K so the statement that
    low temp. ignition delays are overpredicted is not completely accurate.
    %
\end{reviewer}

\begin{response}
    %
    In this context, we are defining low temperature ignition delays as those to the right of the
    experimental NTC region on the Arrhenius plot. We have added the following text near line 308:

    \begin{quote}
        %
        ``... the RMG model tends to over-predict the low-temperature overall ignition delays (i.e.,
        those to the right of the experimental NTC region on the Arrhenius plot)...''
        %
    \end{quote}
    %
\end{response}

@@

\begin{reviewer}
    %
    The measurement of Methyl Valerate at these temperature and pressure conditions constitutes
    important work and the manuscript of well laid out and has potential to be widely circulated.
    The arguments in the article are sound and the figures are clear.
    %
\end{reviewer}

\begin{response}
    Thank you for your comments. A detailed response is below.
\end{response}

\begin{reviewer}
    %
    I would like to ask the authors to check one item definitively before the article is published.
    I am concerned that at these higher pressures and equivalence ratios close to 1, there would be
    ignition which occurs during the compression stroke. This would not be noticed if some of the
    fuel condenses during the charging process and thereby you are dealing with lean mixtures than
    those cited in the paper. The authors are encouraged to make sure this is not the case by
    possibly sampling the charge after filling through a GC or by a quantitative absorption method
    as that used routinely by Stanford.
    %
\end{reviewer}

\begin{response}
    %
    With regards to ignition during the compression stroke, we have compared each reactive pressure
    trace to its corresponding non-reactive trace and found they are in good agreement from the
    beginning to the end of compression, indicating that there is no heat release during the
    compression stroke. With regards to the filling process, we have thoroughly checked the
    temperatures throughout the apparatus and found them to be quite consistent, including the
    mixing tanks, the piping connections, and the reaction chamber itself. Moreover, we have
    conducted several studies in the past utilizing GC to determine mole fractions of components.
    All of these studies determined that the vaporization procedure used on this apparatus is
    sufficient to produce a homogeneous mixture of fuel and oxidizer, even for fuels with relatively
    low vapor pressures. We have added references to these prior studies in the present manuscript
    near line 105:

@@

    \begin{quote}
        %
        ``Previous work has shown this procedure to completely vaporize the fuel and prevent fuel
        cracking during the heating process \cite{Weber2011a,Kumar2009,Das2012}.''
        %
    \end{quote}
    %
\end{response}

\begin{reviewer}
    %
    In the future, a possible direct charging method might be encouraged instead of a mix vessel.
    %
\end{reviewer}

\begin{response}
    %
    We first note that direct charging methods have their own set of disadvantages, and are still
    required to address the issues of fuel condensation and fuel-oxidizer mixing. However, we feel
    that adding a detailed discussion of the advantages and disadvantages of each mixing method to
    the manuscript would be beyond the scope of the current work, and we have not made any changes
    in response to this comment.
    %
\end{response}

### 2017-09-25 revision-1/response/response.tex

\begin{response}
    %
    We have increased the mark size on \cref{fig:simulation-comparison} to better distinguish the
    overall ignition delay and the first-stage ignition delay. On \cref{fig:ign-delay-def}, we made
    several changes to improve the clarity, including swapping the colors of reactive and
    non-reactive pressure traces, changing the non-reactive pressure to a dashed line style, and
    making the black lines thinner.
    %
\end{response}

\begin{reviewer}
    %
    Regarding the statement "The NTC region of MV is mapped out to provide additional information on
    the fidelity of using MV as a biodiesel surrogate." I am not convinced that MV could be a
    biodiesel surrogate, in my opinion it helps in understanding the chemistry of large esters.
    %
\end{reviewer}

\begin{response}
    %
    We agree with the reviewer that it is possible that MV will not be a good surrogate for
    biodiesel. Nonetheless, to our knowledge, there has been no study to suggest that it is not a
    good surrogate, probably because very few studies of MV have been conducted, so there is not
    enough information to judge. However, we think that judging the suitability (or lack thereof)
    for MV as a surrogate of biodiesel is outside the scope of this work, so we are simply noting
    that this study provides part of the necessary experimental information so that others may judge
    for themselves. We also agree with the reviewer that the scope of this work is actually wider
    than we had noted previously, so we have added a note that this study provides further
    information on the autoignition chemistry of large methyl esters near line 41:

    \begin{quote}
        %
        ``The NTC region of MV is mapped out to provide additional information on the fidelity of
        using MV as a biodiesel surrogate and insights into the autoignition chemistry of large
        methyl esters.''
        %
    \end{quote}
    %
\end{response}

@@

\begin{reviewer}
    %
    The authors report an ignition delay of 100 ms at 640 K f=2, MV/air mixture at 15 bar. So at 30
    bar I would expect an ignition delay which is for sure will be more than 10 ms. So I am
    surprised by the statement "For the phi = 2:0 condition, only  PC = 15 bar is considered because
    we could not achieve TC values low enough that the ignition during the compression stroke can be
    prevented'. Can the authors comment on this.
    %
\end{reviewer}

\begin{response}
    %
    The reason for this is because of physical limitations of the experimental apparatus and the
    fuel. In particular, achieving sufficiently low temperatures on our apparatus requires
    increasing the reaction chamber volume at the end of compression, which in turn requires
    substantially higher initial pressures to reach the EOC pressure of \SI{30}{\bar}. The initial
    pressure in the RCM reaction chamber is limited by the pressure in the mixing tank, and the
    total pressure in the mixing tank is limited by the vapor pressure of MV at the preheat
    temperature. The combination of these factors meant that we could not prepare a mixture with
    sufficiently high tank pressure to enable repeated experiments with the EOC volume required to
    reach temperatures where the ignition delay would be measurable. However, we feel that this
    explanation would only confuse the manuscript, so we omit it.
    %
\end{response}

\begin{reviewer}
    %
    The comparison of ignition delays of MV from current work and that of Hadj-Ali et al .[9] would
    be helpful.
    %
\end{reviewer}

\begin{response}
    %
    The ignition delays of MV measured in the work of \citet{Hadj-Ali2009} were conducted at several
    pressures and roughly a single EOC temperature (near \SI{815}{\K}). However, none of the
    pressures considered in that study match the pressures considered in our study, so we feel that
    including the data on one of our existing plots would be very confusing. Moreover, the influence
    of facility-specific effects make a direct comparison of ignition delays from different RCMs
    somewhat challenging, particularly without experimental pressure traces of MV from the work of
    \citet{Hadj-Ali2009} to compare the facility effects. Therefore, we have not added a figure with
    this comparison.
    %
\end{response}

@@

\begin{response}
    %
    We agree with the reviewer that the location of the NTC region with respect to inverse
    temperature is not well captured by the model. It does seem as though the NTC has been shifted
    to higher temperatures with respect to the data, in addition to the model being too reactive.
    This is supported by the newly added \cref{fig:conv-comparison}, which compares adiabatic,
    constant volume simulations with the RCM simulations from \cref{fig:simulation-comparison}. It
    is apparent from this figure that the predicted ignition delays in the $P_C = \SI{15}{\bar}$,
    $\phi = \num{2.0}$ and the $P_C = \SI{30}{\bar}$, $\phi = \num{1.0}$ simulations fall on the low
    temperature side of the NTC, while the $P_C = \SI{15}{\bar}$, $\phi =$ \numlist{0.25;0.50}
    simulated ignition delays are on the high-temperature side of the NTC, but approaching the NTC
    range such that they start to assume the characteristic curvature as the temperature decreases.
    In addition to the figure, we have added some text near line 290 to clarify this point.
    %
\end{response}

\begin{reviewer}
    %
    Figure 5 caption could also be revised.
    %
\end{reviewer}

\begin{response}
    %
    We have modified the text of the caption of \cref{fig:simulation-comparison} to read:

    \begin{quote}
        %
        ``Comparison of experimental (\(\tau\) and \(\tau_1\)) and simulated (\(\tau\)) ignition delays
        computed using the procedure described in \cref{sec:experimental-modeling}. a)
        \SI{15}{\bar}, b) \SI{30}{\bar}.''
        %
    \end{quote}
    %
\end{response}

### 2017-09-25 revision-1/revised-paper/methyl-valerate.tex
Start Jackie's revisions

This work provides additional data for the autoignition of MV. Data is collected in a RCM under
engine relevant conditions spanning from \SIrange{15}{30}{\bar}, equivalence ratios \((\phi)\) from
\numrange{0.25}{2.0}, and temperatures from \SIrange{680}{1050}{\K}. The NTC region of MV is mapped
out to provide additional information on the fidelity of using MV as a biodiesel surrogate and on
the autoignition chemistry of large methyl esters.

@@

\begin{center}
    \captionsetup{type=figure}
    \resizebox{\textwidth}{!}{\input{figures/simulation-comparison.pgf}}
    \caption{Comparison of experimental and simulated ignition delays computed using the procedure
    described in \cref{sec:experimental-modeling}. a) \SI{15}{\bar}, b) \SI{30}{\bar}.}
    \label{fig:simulation-comparison}
\end{center}

@@

To elucidate the underlying reasons for the disagreement between the \citet{Dievart2013} model and
the data, we constructed an additional model using RMG (see
\cref{sec:reaction-mechanism-generator}). As can be seen in \cref{fig:simulation-comparison}a, the
agreement between the RMG model (dashed lines) and the experimental data is similar to the
\citet{Dievart2013} model for the \SI{15}{\bar}, \(\phi =\) \numlist{0.25; 0.5} data. Moreover, the
RMG model predicts a similar NTC region as temperature is decreasing. For the \SI{15}{\bar}, \(\phi =
2.0\) data, the RMG model tends to over-predict the low-temperature ignition delays (i.e., those to
the right of the experimental NTC region on the Arrhenius plot), and does not predict the NTC region
found experimentally. As before, the trend at \SI{30}{\bar}, \(\phi=1.0\) is similar to the
\SI{15}{\bar}, \(\phi=2.0\) data; the RMG model over-predicts the low-temperature ignition delays
and does not predict the experimental NTC region. Finally, as in the \citet{Dievart2013} model,
two-stage ignition is predicted for all of the \(\phi=1.0\) and \(\phi=2.0\) data shown in
\cref{fig:simulation-comparison}. However, the first-stage ignition delays are
\SIrange{0.1}{0.5}{\ms} less than the overall ignition delays, and are not shown on
\cref{fig:simulation-comparison} because they are nearly indistinguishable from the overall ignition
delay.

>>>>

This work provides additional data for the autoignition of MV. Data is collected in a RCM under
engine relevant conditions spanning from \SIrange{15}{30}{\bar}, equivalence ratios \((\phi)\) from
\numrange{0.25}{2.0}, and temperatures from \SIrange{680}{1050}{\K}. The NTC region of MV is mapped
out to provide additional information on the fidelity of using MV as a biodiesel surrogate and
insights into the autoignition chemistry of large methyl esters.

@@

\begin{center}
    \captionsetup{type=figure}
    \resizebox{\textwidth}{!}{\input{figures/simulation-comparison.pgf}}
    \caption{Comparison of experimental (\(\tau\) and \(\tau_1\)) and simulated (\(\tau\)) ignition
    delays computed using the procedure described in \cref{sec:experimental-modeling}. a)
    \SI{15}{\bar}, b) \SI{30}{\bar}.}
    \label{fig:simulation-comparison}
\end{center}

@@

To elucidate the underlying reasons for the disagreement between the \citet{Dievart2013} model and
the data, we constructed an additional model using RMG (see
\cref{sec:reaction-mechanism-generator}). As can be seen in \cref{fig:simulation-comparison}a, the
agreement between the RMG model (dashed lines) and the experimental data is similar to the
\citet{Dievart2013} model for the \SI{15}{\bar}, \(\phi =\) \numlist{0.25; 0.5} data. Moreover, the
RMG model predicts a similar NTC region as temperature is decreasing. For the \SI{15}{\bar}, \(\phi =
2.0\) data, the RMG model tends to over-predict the low-temperature overall ignition delays (i.e.,
those to the right of the experimental NTC region on the Arrhenius plot), and does not predict the
NTC region found experimentally. As before, the trend at \SI{30}{\bar}, \(\phi=1.0\) is similar to
the \SI{15}{\bar}, \(\phi=2.0\) data; the RMG model over-predicts the low-temperature ignition
delays and does not predict the experimental NTC region. Finally, as in the \citet{Dievart2013}
model, two-stage ignition is predicted for all of the \(\phi=1.0\) and \(\phi=2.0\) data shown in
\cref{fig:simulation-comparison}. However, the first-stage ignition delays are
\SIrange{0.1}{0.5}{\ms} less than the overall ignition delays, and are not shown on
\cref{fig:simulation-comparison} because they are nearly indistinguishable from the overall ignition
delay.
